\newglossaryentry{M}{
    name={$M$},
    description={Manifold, mostly 4-d spacetime.}}
    
\newglossaryentry{space mfd}{
    name={$\Sigma$},
    description={3-d space manifold.}}
    
\newglossaryentry{automorphism}{
    name={$\mathrm{Aut}(X)$},
    description={Automorphism group of $X$.}}
    
\newglossaryentry{Lie Alg}{
    name={$\mathrm{Lie}(\mathcal{G})$},
    description={Lie algebra of a Lie group $\mathcal{G}$.}}
    
\newglossaryentry{Diffeo}{
    name={$\mathrm{Diff}(M)$},
    description={Diffeomorphism group of $M$.}}

\newglossaryentry{x}{
    name={$x^a$},
    description={Base space coordinates, mostly spacetime coordinates.}}
    
\newglossaryentry{STInds}{
    name={$a,b,c,...$},
    description={Spacetime indices that range from 0 to 3.}}
    
\newglossaryentry{spatial inds}{
    name={$\alpha, \beta, \gamma,...$},
    description={Spatial indices running from 1 to 3.}}

\newglossaryentry{I}{
    name={$I,J,K,L,...$},
    description={Indices labeling symmetric pairs of spacetime indices running from 0 to 9.}}
    
\newglossaryentry{I_k}{
    name={$I_k,J_k,K_k,L_k,...$},
    description={Higher-order derivative indices running from 0 to $\binom{b+k-1}{n-1}-1$, where $n$ is the base space dimension.}}

\newglossaryentry{additional derivatives}{
    name={$\mathcal{I}_k,\mathcal{J}_k,\mathcal{K}_k,\mathcal{L}_k,...$},
    description={Additional Higher-order derivative indices.}}
    
\newglossaryentry{A}{
    name={$A,B,C,D,...$},
    description={Field bundle fiber indices.}}
    
\newglossaryentry{A'}{
    name={$\tilde{A}, \tilde{B}, \tilde{C}, \tilde{D},... $},
    description={Fiber indices for additional field bundle.}}
    
\newglossaryentry{F}{
    name={$(F, \pi_F, M)$},
    description={Smooth manifold bundle, mostly the field bundle.}}
    
\newglossaryentry{F'}{
    name={$F$},
    description={Field bundle.}}
    
\newglossaryentry{field coordinates}{
    name={$v_A, u_A,...$},
    description={Fiber coordinates, mostly of the field bundle.}}
    
\newglossaryentry{dual field coordinates}{
    name={$v^A, u^A,...$},
    description={Dual coordinates to $v_A, u_A,...$.}}
    
\newglossaryentry{f, g, h,...}{
    name={$f, g, h,...$},
    description={Smooth maps.}}
    
\newglossaryentry{dual}{
    name={$V^{\ast}$},
    description={Dual of an object $V$ (also pull-back of a smooth map).}}

\newglossaryentry{push forward}{
    name={$f_{\ast}$},
    description={Push forward of a smooth map (also lift of a diffeomorphism to any tensor bundle).}}
    
\newglossaryentry{tensor}{
    name={$T^m_nM$},
    description={Rank $(m,n)$ tensor bundle over $M$.}}
    
\newglossaryentry{Sym Tensors}{
    name={$S^m_nM$},
    description={$T^m_nM$ subbundle of fully symmetric tensors.}}
    
\newglossaryentry{forms}{
    name={$\Lambda^nM$},
    description={Bundle of totally anti symmetric rank $(0,n)$ tensors over $M$.}}
    
\newglossaryentry{ASym Tensors}{
    name={$A^m_nM$},
    description={$T^m_nM$ subbundle of fully anti-symmetric tensors.}}
    
\newglossaryentry{tensor prod}{
    name={$\otimes$},
    description={Tensor product.}}
    
\newglossaryentry{wedge}{
    name={$\wedge$},
    description={Wedge product of forms.}}
    
\newglossaryentry{direct sum}{
    name={$\oplus$},
    description={Direct sum of vector spaces or Whitney sum of vector bundles.}}
    
\newglossaryentry{sections}{
    name={$\Gamma(F)$},
    description={Space of sections of a bundle $F$.}}
    
\newglossaryentry{section}{
    name={$G$},
    description={A section of some bundle $F$.}}
    
\newglossaryentry{Jet}{
    name={$j_p^qG$},
    description={1-jet of a section at $p$.}}
    
\newglossaryentry{jet bundle}{
    name={$J^qF$},
    description={q-th order Jet bundle over $F$.}}
    
\newglossaryentry{prolongation}{
    name={$j^q$},
    description={Jet prolongation map.}}
    
\newglossaryentry{jet projection}{
    name={$\pi_{i,j},\pi_i,...$},
    description={Jet bundle projections.}}
    
\newglossaryentry{lift of VF}{
    name={$\xi_F$},
    description={Vector field lifted from $TM$ to $TF$.}}
    
\newglossaryentry{prolonged lifted VF}{
    name={$\xi_{J^qF}$},
    description={Vector field lifted from $TM$ to $T(J^qF)$.}}
    
\newglossaryentry{derivative coords}{
    name={$v_{Ap},v_{AI}, v_{AI_k}$},
    description={Derivative fiber coordinates.}}
    
\newglossaryentry{partial derivative}{
    name={$\frac{\partial}{\partial x^i}, \partial_i$},
    description={Partial derivative or tangential vector.}}
    
\newglossaryentry{partial notation }{
    name={$f^{:a}$},
    description={Short notation for $\partial_a f$.}}

\newglossaryentry{LieDer}{
    name={$\mathcal{L}_{\xi}$},
    description={Lie derivative along $\xi$.}}
    
\newglossaryentry{total derivative from}{
    name={$\mathrm{d}$},
    description={Total derivative of forms.}}
    
\newglossaryentry{total derivative Jet}{
    name={$D_i$},
    description={Jet bundle total derivative (sometimes also called formal derivative).}}
    
\newglossaryentry{variational}{
    name={$\frac{\delta}{\delta v_A}$},
    description={Jet bundle variational derivative.}}
    
\newglossaryentry{volume form valued maps}{
    name={$\mathcal{S}$},
    description={Volume form valued map.}}
    
\newglossaryentry{coord repres}{
    name={$S$},
    description={Corresponding coordinate representation $\mathcal{S} = S \mathrm{d}^nx$.}}

\newglossaryentry{coord volume form}{
    name={$\mathrm{d}^nx$},
    description={Coordinate induced volume form $\mathrm{d}^nx = \overset{n}{\underset{i = 0}{\Lambda}} \mathrm{d}x^i$ .}}

\newglossaryentry{commutator}{
    name={$[.,.]$},
    description={Commutator of vector fields.}}
    
\newglossaryentry{Poisson bracket}{
    name={$\{.,.\}$},
    description={Poisson bracket.}}
    
\newglossaryentry{metric}{
    name={$g_{ab}$},
    description={Metric tensor field.}}
    
\newglossaryentry{3-metric}{
    name={$(g_3)_{\alpha \beta}$},
    description={Slicing induced 3-metric.}}
    
    
\newglossaryentry{Area Metric}{
    name={$G_{abcd}$},
    description={Area Metric tensor field.}}
    
\newglossaryentry{inter}{
    name={$I^A_{abcd}, I^I_{pq}, J_A^{abcd}, J_I^{pq},...$},
    description={Intertwiner.}}
    
\newglossaryentry{multiplicity}{
    name={$\sigma(A)$},
    description={Multiplicity of an index class $A$.}}
    
\newglossaryentry{Gotay-Marsden}{
    name={$C^{Am}_{Bn}$},
    description={Constant component tensor that encodes the Lie algebra action of $\mathrm{Diff}(M)$ on the fibers of a bundle over $M$.}}

\newglossaryentry{time VF}{
    name={$\frac{\partial}{\partial t}$},
    description={Time vector field.}}
    
\newglossaryentry{space at t0}{
    name={$\sigma_{t_0}$},
    description={Embedded space hypersurface at time $t_0$.}}
    
    
\newglossaryentry{tangential}{
    name={$\mathcal{T}\Sigma$},
    description={Tangential subbundle of direct sum split of $TM$.}}
    
\newglossaryentry{tangential Vec}{
    name={$\xi_{\parallel}$},
    description={Tangential part of vector.}}
    
\newglossaryentry{vertical}{
    name={$\mathcal{V}\Sigma, \mathcal{V}F$},
    description={Vertical subbundle of direct sum split of $TM$, or general vertical subbundle.}}
    
\newglossaryentry{vertical Vec}{
    name={$\xi_{\perp}$},
    description={Vertical part of vector.}}
    
\newglossaryentry{tangential def}{
    name={$\mathcal{D}(N^{\alpha})$},
    description={Tangential deformation operator.}}
    
\newglossaryentry{vertical def}{
    name={$\mathcal{H}(N)$},
    description={Vertical deformation operator.}}
    
\newglossaryentry{principal poly}{
    name={$\mathcal{P}, \mathcal{P}_{mat}, \mathcal{P}_{grav},...$},
    description={Principal Polynomial.}}
    
\newglossaryentry{restriction }{
    name={$G_{t_0}$,...},
    description={Restriction to $\Sigma_{t_0}$.}}
    
\newglossaryentry{time derivative }{
    name={$ \dot{G}$},
    description={Lie derivative along $\frac{\partial}{\partial t}$.}}
  
\newglossaryentry{jet decomp }{
    name={$\beta_{t_0} $},
    description={Jet bundle decomposition map.}}
  
\newglossaryentry{time dep Lagrangian }{
    name={$ \mathcal{L}_{t_0}$},
    description={Time dependent Lagrangian .}}
    
\newglossaryentry{embedding }{
    name={$ i_{\Sigma_{t_0}}$},
    description={ Canonical embedding.}}
    
\newglossaryentry{volume valued dual }{
    name={$ (\mathcal{V}F)_{t_0}^{\dagger}$},
    description={...$= \mathcal{V}_{t_0}^{\ast} \otimes \Lambda^3\Sigma_{t_0}$.}}
    
\newglossaryentry{fiber derivative}{
    name={$\mathcal{D}$},
    description={Fiber derivative.}}
    
\newglossaryentry{momenta }{
    name={$\pi^{A}$},
    description={ Conjugate momenta.}}
    
\newglossaryentry{hamiltonian }{
    name={$ \mathcal{H}_{t_0}$},
    description={ Time dependent Hamiltonian.}}
    
\newglossaryentry{configuration bundle }{
    name={$ \mathcal{K}^q_{t_0}$},
    description={ Canonical configuration bundle.}}
    
\newglossaryentry{local phase space function }{
    name={$ \mathcal{f}, \mathcal{g},...$},
    description={ Volume form valued bundle maps on the canonical configuration bundle.}}
    
\newglossaryentry{local phase space functional }{
    name={$ \mathcal{S}_{\mathcal{f}}, \mathcal{S}_{\mathcal{g}}, ...$},
    description={Corresponding functionals.}}
    
\newglossaryentry{weak equality}{
    name={$ \approx$},
    description={ Weak equality.}}
    
\newglossaryentry{primary constraints }{
    name={$ C_{\alpha}, C_0$},
    description={ Primary constraints}}

\newglossaryentry{Ham constraint }{
    name={$\mathbf{H} $},
    description={ Hamiltonian constraint.}}

\newglossaryentry{Diffeomorphism constraint }{
    name={$\mathbf{D}_{\mu} $},
    description={ Diffeomorphism constraint.}}

\newglossaryentry{PDE }{
    name={$R_q $},
    description={ q-th order PDE.}}
    
\newglossaryentry{PDE prolong}{
    name={$R_{q+r} $},
    description={Prolongation to order $q+r$.}}
    
\newglossaryentry{PDE project }{
    name={$R_q^{(1)} $},
    description={ Once prolonged and projected PDE.}}
    
\newglossaryentry{PDE symbol }{
    name={$M_q $},
    description={ Symbol of q-th order PDE.}}
    
\newglossaryentry{Beta }{
    name={$\beta_q(k) $},
    description={ Number of equations of class $k$ in $M_q$.}}
    
\newglossaryentry{Principal Symbol }{
    name={$T^{\Tilde{A}B}(k) $},
    description={ Principal Symbol.}}
    
\newglossaryentry{adjunct matrix }{
    name={$Q_{(A_1...A_n)(B_1...B_n)} $},
    description={ Order $n$ adjunct matrix.}}
    
\newglossaryentry{factor f }{
    name={$f_{(A_1...A_n)(B_1...B_n)}$},
    description={ Factors relating order $n$ adjunct matrix and Principal Polynomial for gauge theories.}}
    
\newglossaryentry{poly at p }{
    name={$\mathcal{P}_p$},
    description={Principal polynoial at $p$.}}
    
\newglossaryentry{poly zero }{
    name={$V_p(\mathcal{P}_p)$},
    description={ Vanishing set of Principal Polynomial at $p$.}}

\newglossaryentry{Hyperbol zero }{
    name={$C_p(\mathcal{P}_p, h)$},
    description={ Hyperbolicity cone of Principal Polynomial at $p$ w.r.t. $h$.}}

\newglossaryentry{Jacoby bracket }{
    name={$\llbracket.,.\rrbracket$},
    description={ Jacobi bracket.}}

\newglossaryentry{collection of all coords }{
    name={$(v_{AI_k})$},
    description={ Tuple of all coordinates up to desired order: $(v_{AI_k}) = (x^m, v_A, v_{Ap}, ...)$.}}
    
\newglossaryentry{expansion eta }{
    name={$(N_{AI_k})$},
    description={ Coordinates of eta induced expansion point.}}

\newglossaryentry{H deviation }{
    name={$(H_{AI_k})$},
    description={Coordinate deviation from expansion point: $H_{AI_k} = (v_{AI_k}) - (N_{AI_k})$ .}}
    
\newglossaryentry{Lagrange jet coordinates}{
    name={$l, l^{A}, ...$ },
    description={ Jet coordinates for equivariance equation.}}

\newglossaryentry{expansion coeffs }{
    name={$a^A, a^{AI}, ...$},
    description={ Expansion coefficients for power series ansatz.}}

\newglossaryentry{Lorentz generators }{
    name={$K^n_{m [rs]}$},
    description={ Generators of Lorentz group.}}
    
\newglossaryentry{expansion general }{
    name={$P_1^A, P_2^{AB}, (T_1^{AB})^C, ...$},
    description={ Expansion coefficients of Principal Polynomial, Principal Symbol, etc.}}

\newglossaryentry{musical iso }{
    name={$\flat$},
    description={Musical isomorphism generated by metric.}}

\newglossaryentry{musical iso area }{
    name={$\flat_{Area}$},
    description={Isomorphism generated by Area Metric.}}



    